Left-linear rules in adhesive categories

### 2024-02-06 icalp/fusione.tex
structure

@@

\begin{remark}In the previous definition, ``decomposition'' corresponds to ``left cancellation'', but we prefer to stick to the name commonly used in literature (see e.g.~\cite{habel2012mathcal}).
\end{remark}
So equipped, we can introduce the notion of \emph{$\mathcal{A}$-Van Kampen square}.

\noindent 
\parbox{10cm}{\begin{definition}[Van Kampen property] Let $\X$ be a category  and consider the two diagrams aside.

\hspace{15pt}Given  a class of arrows $\mathcal{A}\subseteq \mor(\X)$, we say that the bottom square \emph{has the Van Kampen property relatively to $\mathcal{A}$}, or that it  is a \emph{$\mathcal{A}$-Van Kampen square} if:
	\begin{enumerate}
		\item is a pushout square;
		\item 	whenever the cube above has pullbacks as back and left faces and the vertical arrows belong to $\mathcal{A}$, then its top face is a pushout if and only if the front and right faces are pullbacks.
	\end{enumerate}
	Pushout squares which enjoy the ``if'' half of the second point of the condition above are called \emph{$\mathcal{A}$-stable}.

@@

\begin{definition}[$\mathcal{M}$-adhesive category]
	Let $\X$ be a category and consider a subclass $\mathcal{M}$ of the class $\mon(\X)$ of monomorphisms such that:
	\begin{enumerate}
		\item $\mathcal{M}$ contains all isomorphisms and is closed under composition;
		\item $\mathcal{M}$ is stable under pullbacks and pushouts.
	\end{enumerate} 
	We will use $m\colon X\mto Y$ to denote that an arrow $m\colon X\to Y$ belongs to $\mathcal{M}$.

	$\X$ is said to be \emph{$\mathcal{M}$-adhesive} if
	\begin{enumerate}
		\item it has \emph{$\mathcal{M}$-pullbacks}, i.e. pullbacks along arrows of $\mathcal{M}$;
		\item it has \emph{$\mathcal{M}$-pushouts}, i.e. pushouts along arrows of $\mathcal{M}$;
		\item  $\mathcal{M}$-pushouts are $\mathcal{M}$-Van Kampen squares.
	\end{enumerate}

A category $\X$ is said to be \emph{strictly $\mathcal{M}$-adhesive} if $\mathcal{M}$-pushouts are Van Kampen squares.	
\end{definition}

\begin{remark}\label{rem:diff}Our notion of $\mathcal{M}$-adhesivity follows \cite{ehrig2012,ehrig2014adhesive} and is different from the one of \cite{azzi2019essence}. What is called $\mathcal{M}$-adhesivity in that paper corresponds to our strict $\mathcal{M}$-adhesivity. Moreover, in \cite{azzi2019essence} the class $\mathcal{M}$ is assumed to be only stable under pullbacks. However, if $\mathcal{M}$ contains all split monos, then stability under pushouts can be deduced from the other axioms \cite[Prop.~$5.1.21$]{castelnovo2023thesis}.
\end{remark}

\begin{remark}\label{rem:salva} 
	\emph{Adhesivity} and \emph{quasiadhesivity} as defined in \cite{lack2005adhesive,garner2012axioms}  coincide with  strict $\mon(\X) $-adhesivity and strict $\reg(\X)$-adhesivity, respectively. 
\end{remark}

@@

A first result we can prove regards closure under decomposition of $\mathcal{M}$.

\begin{theoremEnd}[category=sec2]{proposition}\label{prop:deco}Let  $\mathcal{A}$ be a class of arrows stable under pullbacks. For every arrow $f\colon X\to Y$ and monomorphism $m\colon Y\to Z$, if $m\circ f \in\mathcal{A}$ then $f\in \mathcal{A}$.
\end{theoremEnd}
\begin{proof}\begin{proofEnd}
	Take the diagram
	\[\xymatrix{X \ar[d]_{\id{X}}\ar[r]^{f}& Y \ar[r]^{\id{Y}}  \ar[d]_{\id{Y}}& Y \ar[d]^{m}\\
		X \ar[r]_{f}& Y \ar[r]_{m} & Z}\]
	Since $m$ is mono the right square is a pullback, while the left square is a pullback by construction. By \Cref{lem:pb1} the whole rectangle is a pullback and the thesis follows.
	\end{proofEnd}
\end{proof}
\begin{corollary}\label{cor:deco}
	In every $\mathcal{M}$-adhesive category $\X$, the class $\mathcal{M}$ is closed under decomposition.
\end{corollary}

Another result which can be immediately established, with the aid of \Cref{prop:pbpo}, is the following one.
\begin{proposition}\label{prop:pbpoad}
	Let $\X$ be an $\mathcal{M}$-adhesive category. Then $\mathcal{M}$-pushouts are pullbacks.
\end{proposition}

@@

\begin{corollary}\label{prop:bal}
If $\X$ is an $\mathcal{M}$-adhesive categories, then every epimorphisms in $\mathcal{M}$ is an isomorphisms. In particular, every adhesive category $\X$ is \emph{balanced}: if a morphism is monic and epic, then it is an isomorphism.
\end{corollary}

$\mathcal{M}$-adhesivity is well-behaved with respect to  the comma construction \cite{mac2013categories}, as shown by the following theorem.
\begin{theorem}[\cite{ehrig2006fundamentals,lack2005adhesive}]\label{lem:comma}
	Let $\A$ and $\B$ be respectively an $\mathcal{M}$-adhesive and an $\mathcal{M}'$-adhesive category. Let also $L:\A\rightarrow \C$ be a functor that preserves $\mathcal{M}$-pushouts, and  $R:\B\rightarrow \C$ be a functor which preserves pullbacks.Then $\comma{L}{R}$ is $\cma{M}{M'}$-adhesive, where 
	\[
	\cma{M}{M}':=\{(h,k)\in \mathcal{A}(\comma{L}{R}) \mid h\in \mathcal{M}, k\in \mathcal{M}'\}\]
\end{theorem}

@@

We end this section recalling two properties of $\mathcal{M}$-adhesive categories:  $\mathcal{M}$-pushout-pullback decomposition and uniqueness of pushouts complements.

@@

	The claim now follows since $k_2$ is a monomorphism.
	\end{proofEnd} 
\end{proof}
\subsection{DPO rewriting systems derivations}
$\mathcal{M}$-adhesive categories are the right context in which to perform abstract rewriting using the so-called ``douple pushout approach'' (DPO). We will recall the basic definitions and properties of this approach to abstract rewriting.

@@

If we look to direct derivations as transitions, it is natural to consider them as edges in a direct graph. Taking objects as vertices objects led us to the following definition \cite{heindel2009category}.

\begin{definition}
	Let $(\X, \R)$ be a left-linear DPO rewrityng system, with $\X$ $\mathcal{M}$-adhesive. The \emph{DPO-derivation graph} of $(\X, \R)$ is the (large)  directed graph $\gpo$ having as vertices the objects of $\X$ and in which an edge between $G$ and $H$ is a direct derivation $\dder{D}\colon G\Mapsto H$.	A \emph{derivation} $\der{D}$ between two objects $G$ and $H$ is a path between them in $\gpo$. The \emph{source} and \emph{target} of $\der{D}$ are, respectively, $G$ and $H$.
\end{definition}

\begin{remark}
We can spell out more explicitly what  a derivation $\dder{D}$ is.  An \emph{empty derivation} starting and ending in $G$ is just $G$ itself.  A \emph{non-empty derivation} $\dder{D}$ is a sequence $\{\dder{D}_i\}_{i=0}^n$ of direct derivations such that:
\begin{enumerate}
	\item for every index $i$, $\dder{D}_i$ is a direct derivation $G_i \Mapsto G_{i+1}$;
	\item $G_0=G$ and $G_{n+1}=H$.
\end{enumerate}

@@

\begin{remark}\label{rem:func}
	Consider a derivation $\der{D}$ in a left-linear DPO rewriting system $(\X, \R)$. We can take the subcategory $\Delta(\der{D})$ of $\X$ given by the arrows appearing in $\der{D}$. This subcategory comes equipped with an inclusion functor $I(\der{D})\colon \Delta(\der{D})\to \X$. Moreover, we can further define $\Deltamin(\der{D})$ as the subcategory of $\Delta(\der{D})$ containing only the bottom row of the derivation.
\end{remark}

\begin{notation}Let $\der{D}=\{\dder{D}_i\}_{i=0}^n$ be a derivation. We will depict the $i^\text{th}$ element $\dder{D}_i$ of $\der{D}$ as in the following diagram.  
	\[\xymatrix{L_i \ar[d]_{m_i}& K_i \ar[d]^{k_i}\ar@{>->}[l]_{l_i} \ar[r]^{r_i} & R_i \ar[d]^{h_i} \\G_{i} & \ar@{>->}[l]^{f_{i}} D_{i} \ar[r]_{g_{i}}& G_{i+1} }\]
	Notice that, in particular, if $\der{D}\colon G\to H$, then $G_0=G$ and $G_{n+1}=H$. When $\der{D}$ has length $1$ we will suppress the indexes. In such case, we will also identify $\der{D}$ with its only element. 
\end{notation}

@@

\begin{proof}\begin{enumerate}
		\item Let us proceed by induction on the length of $\der{D}$.

@@

	\smallskip \noindent$\lgh(\dder{D})=1$. Suppose that $\dder{D}$ has as its single component the derivation
			\[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h} \\G& \ar@{>->}[l]^{f} D \ar[r]_{g}& H  \\}\]
			The arrow $f$ is  the pushout of $l$ and so it is in in $\mathcal{M}$. We can thus consider the $\mathcal{M}$-pushout square
			\[\xymatrix{D \ar@{>->}[d]_{f} \ar[r]^{g} & H \ar@{>->}[d]^{p} \\G \ar[r]_{q}& P }\]
			Since $p\in \mathcal{M}$, the thesis follows immediately from \Cref{rem:cof}.

	\smallskip \noindent$\lgh(\dder{D})\geq 2$. Let $\der{D}$ be $\{\dder{D}_i\}_{i=0}^n$ with $n\geq 1$. Let also $\der{D}'$ be $\{\dder{D}_i\}^{n-1}_{i=0}$ and $\rho_n=(l_n, r_n)$ be the rule applied in $\dder{D}_n$. The pushout of $l_n$ is the arrow $f_n\colon D_n\to G_n$ is in $\mathcal{M}$ and, by inductive hypothesis, $\iota_{G_{n}}\colon G_{n}\to \lpro \der{D}'\rpro$ is in $\mathcal{M}$ too. Thus, we can consider the diagram below, having a pushout as its lower half.
			\[\xymatrix{L_n \ar[d]_{m_{n}}& K_{n} \ar[d]^{k_{n}}\ar@{>->}[l]_{l_{n}} \ar[r]^{r_{n}} & R_{n} \ar[d]^{h_n} \\G_{n} \ar@{>->}[d]_{\iota'_{G_{n}}}& \ar@{>->}[l]^{f_n} D_n \ar[r]_{g_n}& H  \ar@{>->}[d]^{q}\\ \lpro \der{D}' \ar[rr]_{p}\rpro && P}\] 
			Notice that, as in the point above, the arrow $q\colon H\to P$ is the pushout of an element in $\mathcal{M}$, therefore it is enough to show that the diagram so constructed provides a colimiting cocone for $\Delta(\der{D})$.

@@

\todo{a very nice intro}

\begin{definition}[Sequential independence]
  \label{de:sequential-independence}
  Let $(\X, \R)$ be a left-linear DPO rewriting system with $\X$ an $\mathcal{M}$-adhesive category. Let also $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation of length $2$. 	An \emph{independence pair} between $\dder{D}_0$ and $\dder{D}_0$, is a pair of  arrows $i_0\colon R_0\to D_1$ and $i_1\colon L_1\to D_0$ such that the following diagram commutes.
	\[\xymatrix@C=15pt{L_0 \ar[d]_{m_0}&& K_0 \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0 \ar@{.>}@/^.35cm/[drrr]|(.3)\hole_(.4){i_0} \ar[dr]|(.3)\hole_{h_0} && L_1 \ar@{.>}@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1  && \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2}\]
	We will say that $\dder{D}_0$ and $\dder{D}_1$ are \emph{weakly sequentially independent} if an independence pair exists. If such independence pair is unique we will say that $\dder{D}$ and $\dder{D'}$ are \emph{sequentially independent}.
\end{definition}

 Let $(i_0, i_1)$ and $(i'_0, i'_1)$ be independence pairs for the direct derivations $\dder{D}_0$ and $\dder{D}_1$. Notice that, by definition, we have
 \[f_1\circ i_0=h_0=f_1\circ i'_0 \qquad 	g_0\circ i_1=m_1= g_0\circ i'_1\]

	\begin{proposition}\label{rem:weak}
	Let $(\X, \R)$ be a linear DPO rewriting system, then two direct derivations $\dder{D}$ and $\dder{D}'$ are weakly sequential independent if and only if they are sequential independent. 
\end{proposition}

There is another useful properties of weakly sequentially independent direct derivations which is worth mentioning.

\begin{theoremEnd}[category=sec3]{proposition}\label{prop:tec}
	Let $(i_0, i_1)$ be an independence pair between two direct derivations $\dder{D}_0$ and $\dder{D}_1$ in a left-linear DPO rewriting system $(\X,\R)$. Consider the pullback of $f_1\colon D_1\mto G_1$ along $g_0\colon D_0\to G_1$ (first square below),  then there exist the arrows $u_1\colon K_1\to P$ and $u_0\colon K_0\to P$ fitting in the central and right square, Moreover, the right one is always a pullback.
	\[\xymatrix{P \ar@{>->}[r]^{p_0} \ar[d]_{p_1} & D_0 \ar[d]^{g_0}& K_0 \ar[r]^{r_0}  \ar@{.>}[d]_{u_0}& R_0 \ar[d]^{i_0} & K_1 \ar@{>->}[r]^{l_1}  \ar@{.>}[d]_{u_1}& L_1 \ar[d]^{i_1}\\  D_1 \ar@{>->}[r]_{f_1} &G_1 &P \ar[r]_{p_1} & D_1 &P \ar[r]_{p_0}  & D_0}\]
\end{theoremEnd}
\begin{proof}
	\begin{proofEnd}
We start noticing that
\[
f_1\circ i_0\circ r_0=h_0\circ r_0=g_0\circ k_0 \qquad g_0\circ i_1\circ l_1=m_1\circ l_1= f_1 \circ k_1 \] 
Thus there exists the dotted $u_0\colon K_0\to P$, $u_1\colon K_1\to P$  in the diagram
\[\xymatrix{K_0 \ar[d]_{r_0} \ar@{.>}[r]_{u_0}\ar@/^.4cm/[rr]^{k_0} &P \ar@{>->}[r]_{p_0} \ar[d]_{p_1}& D_0 \ar[d]^{g_0}&K_1 \ar@{>->}[d]_{l_1} \ar@{.>}[r]_{u_1}\ar@/^.4cm/[rr]^{k_1} &P \ar[r]_{p_1} \ar@{>->}[d]_{p_0}& D_1 \ar@{>->}[d]^{f_1}\\R_0 \ar@/_.4cm/[rr]_{h_0}\ar[r]^{i_0}& D_1 \ar@{>->}[r]^{f_1}& G_1&L_1 \ar@/_.4cm/[rr]_{m_1}\ar[r]^{i_1}& D_0 \ar[r]^{g_0}& G_1}\]
By \Cref{prop:pbpoad,lem:pb1} we get that the left half of the second rectangle is a pullback. 
	\end{proofEnd} \qedhere 
\end{proof}

We want to be able to to switch the application of two rules used by two consecutive weakly independent direct derivations. This leads us to the following definition.

\begin{definition}[Switch]\label{def:switc}
  \label{de:switch}
 Let $(\X, \R)$ be a left-linear DPO rewriting system with $\X$ an $\mathcal{M}$-adhesive category and suppose that $\der{D}=\{\dder{D}_i\}_{i=0}^1$  is a derivation of length $2$ made by two weakly sequentially independent derivations $\dder{D}_0$, $\dder{D}_1$, with independence pair $(i_0, i_1)$:
 	\[\xymatrix@C=15pt{L_0 \ar[d]_{m_0}&& K_0 \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0 \ar@/^.35cm/[drrr]|(.3)\hole_(.4){i_0} \ar[dr]|(.3)\hole_{h_0} && L_1 \ar@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1  && \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2}\]

@@

\begin{remark}\label{rem:fact}
	Notice that, since $f_{\der{E}, 0}$ is in $\mathcal{M}$, there is at most one $j_1\colon L_0\to D_{\der{E},0}$ such that $m_0=f_{\der{E}, 0} \circ j_1$. Moreover, we already know that there is at most one $j_0\colon R_1\to D_{\der{E},1}$ such that $m_{\der{E},1}=f_{\der{E},1}\circ j_0$, therefore, the independence pair $(j_0,j_1)$ between the component of a switch is uniquely determined.
\end{remark}
\begin{example}\label{ex:diff1} Recall that, given two posets $(P, \leq)$ and $(Q, \leq)$, we can construct the poset $(P, \leq)\oplus(Q,\leq)$ which has as underlying set the disjoint union $P+Q$ and in which every element of $Q$ is greater then every element of $P$, while elements of $P$ and $Q$ are compared using the original orders.

@@

\subsubsection{The $3$-steps lemmas}

As a beginning step in our analysis of switch equivalence, we will establish three lemmas  dealing with derivations of length $3$.

>>>>

% commands for restructuring
\newcommand{\rem}[2]{{\color{blue}#1}{\color{red}#2}}
\renewcommand{\rem}[2]{}

@@

\rem{rimuoverei}{
\begin{remark}
  In the previous definition, ``decomposition'' corresponds to
  ``left cancellation'', but we prefer to stick to the name commonly
  used in literature (see e.g.~\cite{habel2012mathcal}).
\end{remark}
}

So equipped, we can introduce the notion of \emph{$\mathcal{A}$-Van Kampen square}.

\noindent 
\parbox{10cm}{
  \begin{definition}[Van Kampen property]
    Let $\X$ be a category and consider the two diagrams aside.

    \hspace{15pt}
    Given  a class of arrows $\mathcal{A}\subseteq \mor(\X)$, we say that the bottom square
    % \emph{has the Van Kampen property relatively to $\mathcal{A}$}, or that it
    is a \emph{$\mathcal{A}$-Van Kampen square} if:
	\begin{enumerate}
		\item is a pushout square;
		\item 	whenever the cube above has pullbacks as back and left faces and the vertical arrows belong to $\mathcal{A}$, then its top face is a pushout if and only if the front and right faces are pullbacks.
	\end{enumerate}
	Pushout squares which enjoy the ``if'' half of item (2) above are called \emph{$\mathcal{A}$-stable}.

@@

\begin{definition}[$\mathcal{M}$-adhesive category]
  Let $\X$ be a category and let $\mathcal{M}$ be a subclass of
  $\mon(\X)$ 
  % \begin{enumerate}
  % \item $\mathcal{M}$
  including  all isomorphisms, closed under composition,
  % \item $\mathcal{M}$
  and stable under pullbacks and pushouts.
  % \end{enumerate} 
  The category
	%
  $\X$ is said to be \emph{$\mathcal{M}$-adhesive} if
  \begin{enumerate}
  \item it has \emph{$\mathcal{M}$-pullbacks}, i.e. pullbacks along arrows of $\mathcal{M}$;
  \item it has \emph{$\mathcal{M}$-pushouts}, i.e. pushouts along arrows of $\mathcal{M}$;
  \item  $\mathcal{M}$-pushouts are $\mathcal{M}$-Van Kampen squares.
  \end{enumerate}

  A category $\X$ is said to be \emph{strictly $\mathcal{M}$-adhesive}
  if $\mathcal{M}$-pushouts are Van Kampen squares.
\end{definition}
We will use $m\colon X\mto Y$ to denote that an arrow $m\colon X\to Y$ belongs to $\mathcal{M}$.

\rem{rimuoverei}{
\begin{remark}
  \label{rem:diff}
  Our notion of $\mathcal{M}$-adhesivity follows
  \cite{ehrig2012,ehrig2014adhesive} and is different from the one of
  \cite{azzi2019essence}. What is called $\mathcal{M}$-adhesivity in
  that paper corresponds to our strict
  $\mathcal{M}$-adhesivity. Moreover, in \cite{azzi2019essence} the
  class $\mathcal{M}$ is assumed to be only stable under
  pullbacks. However, if $\mathcal{M}$ contains all split monos, then
  stability under pushouts can be deduced from the other
  axioms~\cite[Prop.~$5.1.21$]{castelnovo2023thesis}.
\end{remark}
}

\begin{remark}
  \label{rem:salva}
  \emph{Adhesivity} and \emph{quasiadhesivity} as defined in \cite{lack2005adhesive,garner2012axioms}  coincide with  strict $\mon(\X) $-adhesivity and strict $\reg(\X)$-adhesivity, respectively. 
\end{remark}

@@

\rem{Qui l'idea \`e mostrare che valgono per $\mathcal{M}$ le naturali generalizzazioni delle propriet\`a delle adesive. Credo si possa dirle in prosa e metterle in appendice.}{%
A first result we can prove regards closure under decomposition of $\mathcal{M}$.

\begin{theoremEnd}[category=sec2]{proposition}\label{prop:deco}Let  $\mathcal{A}$ be a class of arrows stable under pullbacks. For every arrow $f\colon X\to Y$ and monomorphism $m\colon Y\to Z$, if $m\circ f \in\mathcal{A}$ then $f\in \mathcal{A}$.
\end{theoremEnd}
\begin{proof}\begin{proofEnd}
	Take the diagram
	\[\xymatrix{X \ar[d]_{\id{X}}\ar[r]^{f}& Y \ar[r]^{\id{Y}}  \ar[d]_{\id{Y}}& Y \ar[d]^{m}\\
		X \ar[r]_{f}& Y \ar[r]_{m} & Z}\]
	Since $m$ is mono the right square is a pullback, while the left square is a pullback by construction. By \Cref{lem:pb1} the whole rectangle is a pullback and the thesis follows.
	\end{proofEnd}
\end{proof}
\begin{corollary}
  \label{cor:deco}
  In every $\mathcal{M}$-adhesive category $\X$, the class $\mathcal{M}$ is closed under decomposition.
\end{corollary}

Another result which can be immediately established, with the aid of \Cref{prop:pbpo}, is the following one.

\begin{proposition}
  \label{prop:pbpoad}
  Let $\X$ be an $\mathcal{M}$-adhesive category. Then
  $\mathcal{M}$-pushouts are pullbacks.
\end{proposition}

@@

\begin{corollary}\label{prop:bal}
If $\X$ is an $\mathcal{M}$-adhesive categories, then every epimorphisms in $\mathcal{M}$ is an isomorphisms. In particular, every adhesive category $\X$ is \emph{balanced}: if a morphism is monic and epic, then it is an isomorphism.
\end{corollary}
}

\todo{P: qui probabilmente si puo' compattare tenendo solo cio' che si usa}
$\mathcal{M}$-adhesivity is well-behaved with respect to  the comma construction \cite{mac2013categories}, as shown by the following theorem.
\begin{theorem}[\cite{ehrig2006fundamentals,lack2005adhesive}]\label{lem:comma}
	Let $\A$ and $\B$ be respectively an $\mathcal{M}$-adhesive and an $\mathcal{M}'$-adhesive category. Let also $L:\A\rightarrow \C$ be a functor that preserves $\mathcal{M}$-pushouts, and  $R:\B\rightarrow \C$ be a functor which preserves pullbacks.Then $\comma{L}{R}$ is $\cma{M}{M'}$-adhesive, where 
	\[
	\cma{M}{M}':=\{(h,k)\in \mathcal{A}(\comma{L}{R}) \mid h\in \mathcal{M}, k\in \mathcal{M}'\}\]
\end{theorem}

@@

\rem{Questo penso tutto in appendice, eventualmente tenendo la decomposition o richiamandola quando si parla di switch}{%
We end this section recalling two properties of $\mathcal{M}$-adhesive categories:  $\mathcal{M}$-pushout-pullback decomposition and uniqueness of pushouts complements.

@@

	The claim now follows since $k_2$ is a monomorphism.
	\end{proofEnd} 
\end{proof}
}
\subsection{DPO rewriting systems derivations}
$\mathcal{M}$-adhesive categories are the right context in which to perform abstract rewriting using the so-called ``douple pushout approach'' (DPO). We will recall the basic definitions and properties of this approach to abstract rewriting.

@@

\rem{Mi pare si possa omettere}{
  If we look to direct derivations as transitions, it is natural to consider them as edges in a direct graph. Taking objects as vertices objects led us to the following definition \cite{heindel2009category}.

\begin{definition}
	Let $(\X, \R)$ be a left-linear DPO rewrityng system, with $\X$ $\mathcal{M}$-adhesive. The \emph{DPO-derivation graph} of $(\X, \R)$ is the (large)  directed graph $\gpo$ having as vertices the objects of $\X$ and in which an edge between $G$ and $H$ is a direct derivation $\dder{D}\colon G\Mapsto H$.	A \emph{derivation} $\der{D}$ between two objects $G$ and $H$ is a path between them in $\gpo$. The \emph{source} and \emph{target} of $\der{D}$ are, respectively, $G$ and $H$.
\end{definition}
}
\begin{remark}
We can spell out more explicitly what  a derivation $\dder{D}$ is.  An \emph{empty derivation} starting and ending in $G$ is just $G$ itself.  A \emph{non-empty derivation} $\dder{D}$ is a sequence $\{\dder{D}_i\}_{i=0}^n$ of direct derivations such that:
\begin{enumerate}
	\item for every index $i$, $\dder{D}_i$ is a direct derivation $G_i \Mapsto G_{i+1}$;
	\item $G_0=G$ and $G_{n+1}=H$.
\end{enumerate}

@@

\begin{remark}
  \label{rem:func}
  Consider a derivation $\der{D}$ in a left-linear DPO rewriting
  system $(\X, \R)$. We can take the subcategory $\Delta(\der{D})$ of
  $\X$ given by the arrows appearing in $\der{D}$. This subcategory
  comes equipped with an inclusion functor
  $I(\der{D})\colon \Delta(\der{D})\to \X$. Moreover, we can further
  define $\Deltamin(\der{D})$ as the subcategory of $\Delta(\der{D})$
  containing only the bottom row of the derivation.
\end{remark}

\begin{notation}
  Let $\der{D}=\{\dder{D}_i\}_{i=0}^n$ be a derivation. We will depict
  the $i^\text{th}$ element $\dder{D}_i$ of $\der{D}$ as in the
  following diagram.
  \[\xymatrix{L_i \ar[d]_{m_i}& K_i \ar[d]^{k_i}\ar@{>->}[l]_{l_i} \ar[r]^{r_i} & R_i \ar[d]^{h_i} \\G_{i} & \ar@{>->}[l]^{f_{i}} D_{i} \ar[r]_{g_{i}}& G_{i+1} }\]
  Notice that, in particular, if $\der{D}\colon G\to H$, then $G_0=G$
  and $G_{n+1}=H$. When $\der{D}$ has length $1$ we will suppress the
  indexes. In such case, we will also identify $\der{D}$ with its only
  element.
\end{notation}

@@

\begin{proof}
  \begin{enumerate}
  \item Let us proceed by induction on the length of $\der{D}$.

@@

    \smallskip \noinden
    t$\lgh(\dder{D})=1$. Suppose that $\dder{D}$
    has as its single component the derivation
    \[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h} \\G& \ar@{>->}[l]^{f} D \ar[r]_{g}& H  \\}\]
    The arrow $f$ is  the pushout of $l$ and so it is in in $\mathcal{M}$. We can thus consider the $\mathcal{M}$-pushout square
    \[\xymatrix{D \ar@{>->}[d]_{f} \ar[r]^{g} & H \ar@{>->}[d]^{p} \\G \ar[r]_{q}& P }\]
    Since $p\in \mathcal{M}$, the thesis follows immediately from \Cref{rem:cof}.

    \smallskip \noindent
    $\lgh(\dder{D})\geq 2$. Let $\der{D}$ be
    $\{\dder{D}_i\}_{i=0}^n$ with $n\geq 1$. Let also $\der{D}'$ be
    $\{\dder{D}_i\}^{n-1}_{i=0}$ and $\rho_n=(l_n, r_n)$ be the rule
    applied in $\dder{D}_n$. The pushout of $l_n$ is the arrow
    $f_n\colon D_n\to G_n$ is in $\mathcal{M}$ and, by inductive
    hypothesis, $\iota_{G_{n}}\colon G_{n}\to \lpro \der{D}'\rpro$ is
    in $\mathcal{M}$ too. Thus, we can consider the diagram below,
    having a pushout as its lower half.
			\[\xymatrix{L_n \ar[d]_{m_{n}}& K_{n} \ar[d]^{k_{n}}\ar@{>->}[l]_{l_{n}} \ar[r]^{r_{n}} & R_{n} \ar[d]^{h_n} \\G_{n} \ar@{>->}[d]_{\iota'_{G_{n}}}& \ar@{>->}[l]^{f_n} D_n \ar[r]_{g_n}& H  \ar@{>->}[d]^{q}\\ \lpro \der{D}' \ar[rr]_{p}\rpro && P}\] 
			Notice that, as in the point above, the arrow $q\colon H\to P$ is the pushout of an element in $\mathcal{M}$, therefore it is enough to show that the diagram so constructed provides a colimiting cocone for $\Delta(\der{D})$.

@@

In this section we discuss the notion of independence between
consecutive rewriting steps, which is a fundamental ingredient of the
rewriting theory, which comes into play when viewing a derivation
sequence as a concurrent computation, the idea being that of
abstracting out the order of execution of independent rewtings. We
observe that the shift to left-linear rules leads to the failure of
various basic properties of the classical notion of independence and
we single out a framework in which these can be re-established.

In the DPO approach to rewriting the canonical notion of independence is that of sequential independence, recalled below.

\begin{definition}[Sequential independence]
  \label{de:sequential-independence}
  Let $(\X, \R)$ be a left-linear DPO rewriting system with $\X$ an
  $\mathcal{M}$-adhesive category. Let also
  $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation of length $2$, with
  $\dder{D}_i$ using rule $\rho_i = (l_i,r_i)$.  An \emph{independence
    pair} between $\dder{D}_0$ and $\dder{D}_1$, is a pair of arrows
  $i_0\colon R_0\to D_1$ and $i_1\colon L_1\to D_0$ such that the
  following diagram commutes.
  \[\xymatrix@C=15pt{L_0 \ar[d]_{m_0}&& K_0 \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0 \ar@{.>}@/^.35cm/[drrr]|(.3)\hole_(.4){i_0} \ar[dr]|(.3)\hole_{h_0} && L_1 \ar@{.>}@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1  && \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2}\]
  We will say that $\dder{D}_0$ and $\dder{D}_1$ are
  % \emph{weakly sequentially independent} if an independence pair exists. If such  independence pair is unique we will say that $\dder{D}$ and $\dder{D'}$ are
  \emph{sequentially independent}.\todo{Eliminata la
    nozione di weakly indepedendent, giusto?}
\end{definition}

Intutitively, the existence of the independence pair is intended to
capture the possibility of switching the application of the two rules:
$f_0 \circ i_1$ is a potential match for $\rho_1$ in $G_0$ and the
existence of $i_0 : R_0 \to D_1$ indicates that $\rho_1$ do not delete
items in the image of $K_0$ and thus $\rho_1$ can be applied after
$\rho_0$.

More precisely, let us formalise what it means for a derivation being the switch of another.

% We want to be able to to switch the application of two rules used by two consecutive weakly independent direct derivations.
% This leads us to the following definition.

\begin{definition}[Switch]
  \label{def:switch}
  Let $(\X, \R)$ be a left-linear DPO rewriting system with $\X$ an
  $\mathcal{M}$-adhesive category and suppose that
  $\der{D}=\{\dder{D}_i\}_{i=0}^1$ is a derivation of length $2$ made
  by two weakly sequentially independent derivations $\dder{D}_0$,
  $\dder{D}_1$, with independence pair $(i_0, i_1)$:
  \[\xymatrix@C=15pt{L_0 \ar[d]_{m_0}&& K_0 \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0 \ar@/^.35cm/[drrr]|(.3)\hole_(.4){i_0} \ar[dr]|(.3)\hole_{h_0} && L_1 \ar@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1  && \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2}\]

@@

The fact that sequential indepedence implies switchability always holds for linear rules. [REF]
The result is so fundamental that it is not even stated, in the sense that switchability is not introduced axiomatically as in defition~\ref{def:switch}, but it is based on the explicit construction of a switch.

When working with left-linear rewriting systems, in general (weakly) sequential independence does not imply switchability (while the converse implication clearly holds), as shown by the example below.

\rem{remove}{
  \begin{remark}
    \label{rem:fact}
    Notice that, since $f_{\der{E}, 0}$ is in $\mathcal{M}$, there is
    at most one $j_1\colon L_0\to D_{\der{E},0}$ such that
    $m_0=f_{\der{E}, 0} \circ j_1$. Moreover, we already know that
    there is at most one $j_0\colon R_1\to D_{\der{E},1}$ such that
    $m_{\der{E},1}=f_{\der{E},1}\circ j_0$, therefore, the
    independence pair $(j_0,j_1)$ between the component of a switch is
    uniquely determined.
\end{remark}
}

\begin{example}
  \label{ex:diff1}
  Recall that, given two posets $(P, \leq)$ and $(Q, \leq)$, we can
  construct the poset $(P, \leq)\oplus(Q,\leq)$ which has as
  underlying set the disjoint union $P+Q$ and in which every element
  of $Q$ is greater then every element of $P$, while elements of $P$
  and $Q$ are compared using the original orders.

@@

\subsection{Switch equivalence via permutation equivalence}

\begin{itemize}
\item valutare se si puo' fare senza switch equivalence, forse meglio con

\item linear: switch iff consistent permutation

\item left-linear: only one implication
\end{itemize}

\subsection{The $3$-steps lemmas}
